---
title: "R for PA Project"
output: html_document
date: "2024-12-02"
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(reshape2)
library(caret)
library(RColorBrewer)
library(stats)
library(cluster)
library(ggplot2)
library(scales)
library(gridExtra)
library(dplyr)
library(knitr)
library(kableExtra)
library(FactoMineR)
library(factoextra)
```

```{r setup, include=FALSE}
# Download the dataset from GitHub
url <- "https://raw.githubusercontent.com/fongbubble/UoB_MGRCM0034_Car_Sales/main/car_sales.csv"
csv_file_path <- tempfile(fileext = ".csv")
download.file(url, destfile = csv_file_path)
print(paste("CSV File Path:", csv_file_path))
```

```{r setup, include=FALSE}
# Read the CSV file
df <- read_csv(csv_file_path)
```

Explore Dataset

```{r}
# Explore Dataset
str(df)
```

Explore Dataset

```{r}
# Explore Dataset
summary(df)
```

Search for NaN

```{r}
# Function to search for NaN and NA values in the dataframe
count_NaN_NA <- function(df) {
  # Initialize lists to store counts
  nan_counts <- list()
  na_counts <- list()
  
  # Loop through each column in the dataframe
  for (col_name in names(df)) {
    # Check if the column is numeric
    nan_count <- sum(is.nan(df[[col_name]]))
    na_count <- sum(is.na(df[[col_name]]))
    
    # Store counts in lists with column names
    nan_counts[[col_name]] <- nan_count
    na_counts[[col_name]] <- na_count
  }
  
  # Combine the counts into a dataframe
  counts_df <- data.frame(
    Column = names(df),
    NaN_Count = unlist(nan_counts),
    NA_Count = unlist(na_counts)
  )
  
  return(counts_df)
}

nan_na_counts <- count_NaN_NA(df)
print(nan_na_counts)
```

Search for Duplicate

```{r}
# Search for Duplicate
sum(duplicated(df))
```

Find Unique Company and Model

```{r}
# Find the number of unique values in the 'Model' column
num_unique_models <- df %>% summarise(Unique_Models = n_distinct(Model))

# Find the number of unique values in the 'Model' column
num_unique_brands <- df %>% summarise(Unique_Brands = n_distinct(Company))

# Print the result
print(num_unique_models)
print(num_unique_brands)
```

Scatter plot of Price and Income

```{r}
# Create the scatter plot
p <- ggplot(df, aes(x = `Annual Income`, y = `Price ($)`)) +
  geom_point(color = '#32CD32') +
  ggtitle('Scatter plot of Annual Income vs. Price ($)') +
  xlab('Annual Income') +
  ylab('Price ($)') +
  scale_x_continuous(labels = comma) +  # Adjust x-axis labels to use commas
  theme_minimal(base_size = 15) +
  theme(
    plot.title = element_text(size = 20, hjust = 0.5),  # Adjust the title font size here
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14)
  )

# Save the plot
ggsave("Annual_Income_vs_Price.png", p, width = 15, height = 7, units = "in")

# Print the plot
print(p)
```

Body Type by Gender

```{r}
# Order Gender levels
df$Gender <- factor(df$Gender, levels = c("Female", "Male"))

# Plot Histogram of Body Style by Gender with Female first
p2 <- ggplot(df, aes(x = `Body Style`, fill = Gender)) +
  geom_histogram(stat = "count", position = "dodge") +
  ggtitle('Body Style by Gender') +
  theme_minimal(base_size = 1) +
  theme(
    plot.title = element_text(size = 14, hjust = 0.5), 
    axis.text = element_text(size = 10), 
    axis.title = element_text(size = 12),
    legend.text = element_text(size = 12), 
    legend.title = element_text(size = 12)  
  )

# Save the figure
ggsave("Gender_vs_Body.png", p2, width = 6, height = 6, units = "in")

print(p2)

```

Price By Region

```{r}
# Choosing Dealer_Region
df$Dealer_Region <- factor(df$Dealer_Region, levels = unique(df$Dealer_Region))

# Analyze how region affects the price of the car using a Boxplot
R <- ggplot(df, aes(x = Dealer_Region, y = `Price ($)`, fill = Dealer_Region, group = Dealer_Region)) +
  geom_boxplot() +
  scale_fill_brewer(palette = "Spectral") +
  labs(x = "Dealer Region", y = "Price ($)", title = "Effect of Dealer Region on Car Price") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_minimal()

# Save the figure
ggsave("Dealer Region vs Price.png", R, width = 10, height = 5, units = "in")

print(R)
```

Descriptive Statistic on Annual Income and Cars Price

```{r}
# Function to calculate mode and its frequency
mode_func <- function(x) {
  ux <- unique(na.omit(x))
  tab <- tabulate(match(x, ux))
  mode_val <- ux[which.max(tab)]
  mode_freq <- max(tab)
  return(c(mode_val, mode_freq))
}

# Function to set numbers format 
format_number <- function(x) {
  format(round(x), big.mark = ",", scientific = FALSE)
}

# Calculate descriptive statistics for Annual Income
annual_income_stats <- df %>%
  summarise(
    Mean = mean(`Annual Income`, na.rm = TRUE),
    Median = median(`Annual Income`, na.rm = TRUE),
    SD = sd(`Annual Income`, na.rm = TRUE),
    Min = min(`Annual Income`, na.rm = TRUE),
    Max = max(`Annual Income`, na.rm = TRUE),
    Mode = as.numeric(mode_func(`Annual Income`)[1]),
    Mode_Frq = as.numeric(mode_func(`Annual Income`)[2])
  ) %>%
  mutate(across(everything(), format_number))

# Calculate descriptive statistics for Price ($)
price_stats <- df %>%
  summarise(
    Mean = mean(`Price ($)`, na.rm = TRUE),
    Median = median(`Price ($)`, na.rm = TRUE),
    SD = sd(`Price ($)`, na.rm = TRUE),
    Min = min(`Price ($)`, na.rm = TRUE),
    Max = max(`Price ($)`, na.rm = TRUE),
    Mode = as.numeric(mode_func(`Price ($)`)[1]),
    Mode_Frq = as.numeric(mode_func(`Price ($)`)[2])
  ) %>%
  mutate(across(everything(), format_number))

# Combine both sets of statistics into one table
combined_stats <- bind_rows(
  Annual_Income = annual_income_stats,
  Price = price_stats,
  .id = "Variable"
)

# Print the combined statistics table
combined_stats %>%
  kable(align = "r") %>%
  kable_styling(full_width = FALSE, position = "center")
```

Gender Plot

```{r}
# Pie Chart for Gender
gender_counts <- df %>% count(Gender)
gender_counts <- gender_counts %>% mutate(perc = n / sum(n) * 100)
G <- ggplot(gender_counts, aes(x = "", y = n, fill = Gender)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c('#ff0054', '#0582ca')) +
  ggtitle('Gender Proportion') +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size = 14, hjust = 0.5), axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label = paste0(round(perc, 1), "%")), position = position_stack(vjust = 0.5), size = 4)

# Save the figure
ggsave("Gender Proportion.png", G, width = 7, height = 5, units = "in")

print(G)
```

Annual Income Distribution

```{r}
# Plot Annual Income Distribution
AD <- ggplot(df, aes(x = `Annual Income`)) +
  geom_histogram(aes(y = after_stat(count)), bins = 295, fill = "#57cc99", alpha = 1, na.rm = TRUE) +
  geom_density(aes(y = after_stat(count)), color = "#57cc99", size = 1, na.rm = TRUE) +
  scale_x_continuous(name = "Annual Income", limits = c(0, 3000000), breaks = seq(0, 11200000, by = 100000)) +
  labs(title = "Annual Income Distribution") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        plot.title = element_text(size = 16), 
        axis.title.x = element_text(size = 13))

# Save the figure
ggsave("Annual Income Distribution.png", AD, width = 12, height = 5, units = "in")

print(AD)
```

Annual Income Distribution Zoom

```{r}
# Plot Annual Income Distribution for specific range
ADz <- ggplot(df, aes(x = `Annual Income`)) +
  geom_histogram(aes(y = after_stat(count)), bins = 60, fill = "#57cc99", alpha = 1, na.rm = TRUE) +
  geom_density(aes(y = after_stat(count)), color = "#57cc99", size = 1, na.rm = TRUE) +
  scale_x_continuous(name = "Annual Income", limits = c(0, 1000000), breaks = seq(0, 11200000, by = 50000)) +
  labs(title = "") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1), 
        plot.title = element_text(size = 16), 
        axis.title.x = element_text(size = 13))

# Save the figure
ggsave("Annual Income Distribution Zoom.png", ADz, width = 12, height = 5, units = "in")

print(ADz)
```

Vehicle Plot

```{r}
# Plot 1: Distribution of Price
p1 <- ggplot(df, aes(x = `Price ($)`)) +
  geom_histogram(bins = 50, fill = '#da2c38', color = '#da2c38', alpha = 0.7) +
  geom_density(aes(y = after_stat(count)), color = 'blue', size = 1) +
  ggtitle('Distribution of Car Price ($)') +
  xlab('Price ($)') +
  ylab('Frequency') +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size = 14, hjust = 0.5), axis.text = element_text(size = 10), axis.title = element_text(size = 12))

# Plot 2: Pie chart for Body Style distribution
body_style_counts <- df %>% count(`Body Style`)
body_style_counts <- body_style_counts %>% mutate(perc = n / sum(n) * 100)
body_style_counts <- body_style_counts %>% arrange(desc(perc))
p2 <- ggplot(body_style_counts, aes(x = "", y = n, fill = `Body Style`)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c('#8ecae6', '#219ebc', '#83c5be', '#ffb703', '#fb8500')) +
  ggtitle('Distribution of Vehicles by Body Style') +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size = 14, hjust = 0.5), axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label = paste0(round(perc, 1), "%")), position = position_stack(vjust = 0.5), size = 4)

# Plot 3: Stacked bar chart for Color and Body Style
color_body_pivot <- df %>%
  count(Color, `Body Style`) %>%
  pivot_wider(names_from = `Body Style`, values_from = n, values_fill = list(n = 0))
color_body_pivot_long <- color_body_pivot %>%
  gather(key = "Body Style", value = "Count", -Color)
p3 <- ggplot(color_body_pivot_long, aes(x = Color, y = Count, fill = `Body Style`)) +
  geom_bar(stat = "identity") +
  ggtitle('Vehicle Colors by Body Style') +
  xlab('Color') + 
  ylab('Number of Vehicles') +
  scale_y_continuous(limits = c(0, 12000)) +
  scale_fill_manual(values = c('#8ecae6', '#219ebc', '#023047', '#ffb703', '#fb8500')) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size = 14, hjust = 0.5), axis.text.x = element_text(angle = 50, hjust = 1, size = 10), axis.text.y = element_text(size = 10), axis.title = element_text(size = 12))

# Plot 4: Pie Chart for Transmission (Auto vs Manual) 
transmission_counts <- df %>% count(Transmission)
transmission_counts <- transmission_counts %>% mutate(perc = n / sum(n) * 100)
p4 <- ggplot(transmission_counts, aes(x = "", y = n, fill = Transmission)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c('#83c5be', 'steelblue')) +
  ggtitle('Transmission Proportion: Auto vs Manual') +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(size = 14, hjust = 0.5), axis.title.x = element_blank(), axis.title.y = element_blank(), axis.text = element_blank(), axis.ticks = element_blank()) +
  geom_text(aes(label = paste0(round(perc, 1), "%")), position = position_stack(vjust = 0.5), size = 4)

# Create a layout for the subplots
layout <- grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2, 
                       widths = unit(c(1, 1.2), "null"), heights = unit(c(1, 1.2), "null"))

# Save the figure
ggsave("vehicle_data_visualizations.png", layout, width = 14, height = 10, units = "in")
```

Price by Body Style

```{r}
# Boxplot of Price by Body Style
p <- ggplot(df, aes(x = `Price ($)`, y = `Body Style`, fill = `Body Style`)) +
  geom_boxplot(width = 0.3) +
  scale_fill_manual(values = c('#8ecae6', '#219ebc', '#06d6a0', '#ffb703', '#fb8500')) +
  ggtitle('Price by Body Style') +
  xlab('Price ($)') +
  ylab('Body Style') +
  theme_minimal(base_size = 8) +
  theme(plot.title = element_text(size = 12, hjust = 0.5), axis.text = element_text(size = 8), axis.title = element_text(size = 12), legend.position = "none")

# Save the figure
ggsave("Price by Body Style.png", p, width = 7, height = 5, units = "in")

print(p)
```

Label Encoder

```{r}
# Define the columns to encode
columns_to_encode <- c('Gender', 'Transmission', 'Body Style', 'Company', 'Model')

# Encode columns using factor levels
for (col in columns_to_encode) {
  df[[paste0(col, '_Encoded')]] <- as.numeric(factor(df[[col]]))
}
```

Standard Scaler for Selected Features

```{r}
# Feature Selection for Scaler
X1 <- df %>% select(`Annual Income`, `Price ($)`, `Company_Encoded`, `Model_Encoded`)

# Standardize the features
scaler <- preProcess(X1, method = c("center", "scale"))
X1_scaled <- predict(scaler, X1)
```

Correlation Plot

```{r}
# Copy the dataframe
encoded_df <- df

# Encode categorical variables using Label Encoding
for(col in names(encoded_df)) {
  if(is.factor(encoded_df[[col]]) || is.character(encoded_df[[col]])) {
    encoded_df[[col]] <- as.numeric(factor(encoded_df[[col]]))
  }
}

# Check the structure of encoded_df
str(encoded_df)

# Calculate correlation matrix
corr_matrix <- cor(encoded_df, use = "pairwise.complete.obs")

# Ensure melt function 
melted_corr <- melt(corr_matrix, na.rm = TRUE)

# Plot the heatmap
Cr <- ggplot(data = melted_corr, aes(x=Var1, y=Var2, fill=value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "#ff595e", high = "#1982c4", mid = "#ffca3a", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name="Correlation") +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, 
                                   size = 12, hjust = 1)) +
  coord_fixed() +
  ggtitle("Correlation Heatmap for All Encoded Variables") +
  theme(plot.title = element_text(hjust = 0.5))

# Save the figure
ggsave("Correlation.png", Cr, width = 15, height = 8, units = "in")

print(Cr)
```

Apply PCA

```{r}
# Apply PCA to reduce dimensionality to 2 components
pca_result <- PCA(X1_scaled, ncp = 2, graph = FALSE)

# Extract the coordinates for the variables (features)
components <- pca_result$var$coord[, 1:2]  # Get coordinates for the first 2 components

# Convert the components to a dataframe and transpose it
components_df <- as.data.frame(components)
rownames(components_df) <- c('Annual Income', 'Price ($)', 'Company_Encoded', 'Model_Encoded')
colnames(components_df) <- c('PC1', 'PC2')

# Print the DataFrame
print(components_df)
```

Feature Contribution

```{r}
# Apply PCA to reduce dimensionality to 2 components
pca_result <- PCA(X1_scaled, ncp = 2, graph = FALSE)

# Extract the coordinates for the variables (features)
components <- pca_result$var$coord[, 1:2]  # Get coordinates for the first 2 components

# Convert the components to a dataframe
components_df <- as.data.frame(components)
colnames(components_df) <- c('PC1', 'PC2')  # Set column names to principal components

# Set row names
rownames(components_df) <- rownames(pca_result$var$coord)

# Convert to long format for plotting
components_df$Features <- rownames(components_df)  # Add Features column for melting
components_long <- melt(components_df, id.vars = 'Features', variable.name = 'Principal_Component', value.name = 'Contribution')

# Bar plot showing feature contribution to each principal component
FC <- ggplot(components_long, aes(x = Principal_Component, y = Contribution, fill = Features)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  scale_fill_brewer(palette = 'Spectral') +
  labs(title = 'Feature Contribution to Each Principal Component', x = 'Principal Components', y = 'Contribution') +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  theme(legend.title = element_text(face = 'bold'), legend.position = 'top')

# Save the figure with the specified size
ggsave("Feature Contribution.png", FC, width = 7, height = 5, units = "in")

print(FC)
```

GridSearch to tune hyperparameter

```{r}
# Extract the PCA-transformed data
X1_pca <- pca_result$ind$coord  # This gives the coordinates in the new PCA space

# Define parameter grid for hyperparameter tuning
param_grid <- expand.grid(n_clusters = 3:9,
                          init = c('k-means++', 'random'),
                          max_iter = c(300, 500))

# Function to run K-Means with different parameters and compute total within-cluster
run_kmeans <- function(data, centers, max_iter, nstart = 10) {
  set.seed(24150)
  kmeans_result <- kmeans(data, centers = centers, iter.max = max_iter, nstart = nstart)
  return(kmeans_result$tot.withinss)
}

# Data frame to store the results
results <- data.frame(n_clusters = numeric(), init = character(), max_iter = numeric(), tot_withinss = numeric())

# Perform grid search
for (i in 1:nrow(param_grid)) {
  k <- param_grid$n_clusters[i]
  init_method <- param_grid$init[i]
  max_iter_val <- param_grid$max_iter[i]
  
  # Run K-Means
  tot_withinss <- run_kmeans(X1_pca, centers = k, max_iter = max_iter_val, nstart = 25)
  
  # Store the results
  results <- rbind(results, data.frame(n_clusters = k, init = init_method, max_iter = max_iter_val, tot_withinss = tot_withinss))
}

# Identify the best parameters based on the minimum total within-cluster sum of squares
best_params <- results[which.min(results$tot_withinss), ]
print(best_params)
```

PCA with K-means clustering

```{r}
# Fit the best K-Means algorithm
set.seed(24150)
best_algorithm <- kmeans(X1_pca, centers = 9, nstart = 25, iter.max = best_params$max_iter)
df$Cluster <- factor(best_algorithm$cluster, levels = 1:9, labels = c(
  "High-Income, Cost-Conscious Buyers",  # Cluster 1
  "Low-Income, Value-Driven Buyers",     # Cluster 2
  "Mid-Income, Practical Buyers",        # Cluster 3
  "Mid-Income, Brand-Focused Buyers",    # Cluster 4
  "High-Income, Luxury Seekers",         # Cluster 5
  "High-Income, Brand-Focused Buyers",   # Cluster 6
  "Mid-Income, Budget-Conscious Buyers", # Cluster 7
  "Low-Income, Budget-Focused Buyers",   # Cluster 8
  "Low-Income, Practical Buyers"         # Cluster 9
))
centroids1 <- as.data.frame(best_algorithm$centers)
colnames(centroids1) <- c('PC1', 'PC2')

# Define the custom cluster order to match the right graph
cluster_order <- c(
  "High-Income, Cost-Conscious Buyers",  # Cluster 1
  "Low-Income, Value-Driven Buyers",     # Cluster 2
  "Mid-Income, Practical Buyers",        # Cluster 3
  "Mid-Income, Brand-Focused Buyers",    # Cluster 4
  "High-Income, Luxury Seekers",         # Cluster 5
  "High-Income, Brand-Focused Buyers",   # Cluster 6
  "Mid-Income, Budget-Conscious Buyers", # Cluster 7
  "Low-Income, Budget-Focused Buyers",   # Cluster 8
  "Low-Income, Practical Buyers"         # Cluster 9
)

# Create grid for visualizing K-Means clustering boundaries
x_min <- min(X1_pca[, 1]) - 1
x_max <- max(X1_pca[, 1]) + 1
y_min <- min(X1_pca[, 2]) - 1
y_max <- max(X1_pca[, 2]) + 1
grid <- expand.grid(x = seq(x_min, x_max, by = 0.05), y = seq(y_min, y_max, by = 0.05))

# Assign cluster labels to the grid points
grid$Cluster <- apply(grid, 1, function(row) {
  dists <- apply(centroids1, 1, function(center) {
    sqrt((row[1] - center[1])^2 + (row[2] - center[2])^2)
  })
  which.min(dists)
})

# Convert Cluster to factor with custom labels for plotting
grid$Cluster <- factor(grid$Cluster, levels = 1:9, labels = c(
  "High-Income, Cost-Conscious Buyers",  # Cluster 1
  "Low-Income, Value-Driven Buyers",     # Cluster 2
  "Mid-Income, Practical Buyers",        # Cluster 3
  "Mid-Income, Brand-Focused Buyers",    # Cluster 4
  "High-Income, Luxury Seekers",         # Cluster 5
  "High-Income, Brand-Focused Buyers",   # Cluster 6
  "Mid-Income, Budget-Conscious Buyers", # Cluster 7
  "Low-Income, Budget-Focused Buyers",   # Cluster 8
  "Low-Income, Practical Buyers"         # Cluster 9
))

# Plot PCA scatter plot with K-Means clustering
SP <- ggplot(df, aes(x = X1_pca[, 1], y = X1_pca[, 2], color = Cluster)) +
  geom_point(alpha = 0.8) +
  geom_point(data = centroids1, aes(x = PC1, y = PC2), color = 'red', shape = 8, size = 4) +
  scale_color_manual(values = RColorBrewer::brewer.pal(9, "Spectral"), labels = cluster_order) +
  labs(title = 'PCA Scatter Plot with K-Means Clustering', x = 'PC1: Car Preference & Affordability Component', y = 'PC2: Income Component') +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 15, color = 'navy'),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 10),
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 12),
    legend.key.size = unit(1.2, "lines")
  ) +
  geom_tile(data = grid, aes(x = x, y = y, fill = Cluster), alpha = 0.3, inherit.aes = FALSE) +
  scale_fill_manual(values = RColorBrewer::brewer.pal(9, "Spectral"), labels = cluster_order)

# Save the figure
ggsave("PCA Scatter Plot.png", SP, width = 10, height = 5, units = "in")

print(SP)
```
